---
title: "EOSC573 data analysis"
output: r_notebook
---


```{r}
library(tidyverse)
library(ggplot2)
library(purrr)
library(RADanalysis)
library(vegan)
library(patchwork)
library(ggthemes)
library(RColorBrewer)
library(marmap)
library(rgdal)
library(raster)
library(ggOceanMapsData)
library(ggOceanMaps)
library(viridis)
```

**UPDATED AS OF APR 23 (notes below)**
- Changed colour palette and coerced graph fill using fct_inorder(), new plots have a "2" suffix.
- Biomass contribution by size class
- Bathymetry & map stuff

(Methods) Table: Key features
(Methods) Image: Sample sites highlighted on map

Abundance graphs
  Species
  Classes

Relative composition stacked bar graphs
  Species
  Classes
  
Biomass graphs
  Biomass between sites overall
  Biomass % composition stacked bars
  Biomass large (5+mm) vs med (2-5mm) vs small (0-2mm) between groups

Diversity stuff (table)
  Shannon Div Index for each site
  Species evenness for each site

Plus size histograms

Biomass contribution by size

Bathymetry via marmap package

```{r}
#Abundance and species comp
densitiessites <- read.csv("~/Desktop/densitysites.csv", row.names = 1)
densitiessites2<- tibble::rownames_to_column(densitiessites, "Site")

density_plot2 <- densitiessites2%>%
  ggplot(aes(x = Site, y = Average.Density))+
  geom_bar(stat = 'identity', colour = "steelblue", fill = "steelblue")+
  labs(x = "Sites in Barkley Sound",
       y = "Zooplankton Density (ind/m^3)",
       title = "Zooplankton Abundance in Barkley Sound")+
  theme(plot.title = element_text(hjust = 0.5))+
  scale_x_discrete(limits=rev)

density_plot2 <- density_plot2 +
  theme(axis.title = element_text(size = 18)) +
  theme(plot.title = element_text(size = 20)) +
  theme(axis.text = element_text(size = 18))
  #theme(legend.title = element_text(size = 18)) +
  #theme(legend.text = element_text(size = 16))
density_plot2

ggsave(here::here("density_plot2.png"), height = 9, width = 12)


# species comp
speciespercent <- read.csv("./data/speciescomposition2.csv", row.names = 1)

speciespercent2 <- tibble::rownames_to_column(speciespercent, "Species name")
speciespercent2 <- speciespercent2 %>%
  pivot_longer(!`Species name`, names_to = "Site", values_to = "Count")

#Custom colour palette

#mycolours = c(brewer.pal(name="YlOrRd", n=9), "#8C510A", "#993404", "#662506", "#252525", "#000000", "#1B9E77", "#7570B3", "#E7298A", "#1D91C0", "#7FBC41", "#4D9221", "#276419", "#E31A1C", "#FFFF99", "#FDBF6F", "#969696", "#1F78B4")
mycolours = c(viridis(n = 14, option = "D"), viridis(n = 12, direction = 1, option = "A"))

speciescomp_plot <- speciespercent2 %>%
  ggplot(aes(fill = fct_inorder(`Species name`), x = Site, y = Count)) + 
    geom_bar(position="fill", stat="identity")+
  labs(x = "Site",
       y = "Proportion of Sample",
       fill = "Species Name",
       title = "Species Composition of Zooplankton in Barkley Sound")+
  theme(plot.title = element_text(hjust = 0.5))+
  #scale_fill_excel_new("Office Theme")+
  #scale_fill_viridis(discrete = TRUE, option = "C")+
  scale_x_discrete(limits=rev)+
  scale_fill_manual(values = mycolours)

speciescomp_plot <- speciescomp_plot +
  theme(axis.title = element_text(size = 18)) +
  theme(plot.title = element_text(size = 20)) +
  theme(axis.text = element_text(size = 18)) +
  theme(legend.title = element_text(size = 18)) +
  theme(legend.text = element_text(size = 16))
print(speciescomp_plot)

ggsave("speciescomp_plot2.png", height = 9, width = 14)

#species comp by group

compositiongroups <- read.csv("~/Desktop/compositiongorups.csv", row.names = 1)

compositiongroups2 <- tibble::rownames_to_column(compositiongroups, "Taxon")
compositiongroups2$N1 <- as.double(compositiongroups2$N1)
compositiongroups2 <- compositiongroups2 %>%
  pivot_longer(!`Taxon`, names_to = "Site", values_to = "Count")


groupcomp_plot <- compositiongroups2 %>%
  ggplot(aes(fill = `Taxon`, x = Site, y = Count)) + 
    geom_bar(position="fill", stat="identity")+
  labs(x = "Site",
       y = "Proportion of Sample",
       title = "Taxon Composition of Zooplankton in Barkley Sound")+
  theme(plot.title = element_text(hjust = 0.5))+
  #scale_fill_excel_new("Office Theme")+
  #scale_fill_viridis(discrete = TRUE, option = "C")+
  scale_x_discrete(limits=rev)+
  scale_fill_manual(values = mycolours)

groupcomp_plot <- groupcomp_plot +
  theme(axis.title = element_text(size = 18)) +
  theme(plot.title = element_text(size = 20)) +
  theme(axis.text = element_text(size = 18)) +
  theme(legend.title = element_text(size = 18)) +
  theme(legend.text = element_text(size = 16)) +
  theme(plot.subtitle = element_text(size = 16))
groupcomp_plot

ggsave(here::here("groupcomp_plot.png"), height = 9, width = 12)


```


```{r}
#Size histograms
sizebins <- read.csv("~/Desktop/sizes.csv")

srsize_plot <- sizebins %>%
  ggplot(aes(x = X , y = SR))+
  geom_bar(stat = 'identity', position = 'dodge', colour = "steelblue", fill = "steelblue")+
  labs(x = 'Body Length Size Classes',
       y = "Proportion of Individuals (%)",
       title = 'Site SR1 at Sarita River',
       subtitle = '30m tow depth'
       )+
  theme(plot.title = element_text(hjust = 0.5), plot.subtitle = element_text(hjust = 0.5))

srsize_plot <- srsize_plot +
  theme(axis.title = element_text(size = 18)) +
  theme(plot.title = element_text(size = 20)) +
  theme(axis.text = element_text(size = 18)) +
  theme(legend.title = element_text(size = 18)) +
  theme(legend.text = element_text(size = 16)) +
  theme(plot.subtitle = element_text(size = 16))
view(srsize_plot)

n1size_plot <- sizebins %>%
  ggplot(aes(x = X , y = N1))+
  geom_bar(stat = 'identity', position = 'dodge', colour = "steelblue", fill = "steelblue")+
  labs(x = 'Body Length Size Classes',
       y = "Proportion of Individuals (%)",
       title = 'Site N1 in Trevor Channel',
       subtitle = '30m tow depth'
       )+
  theme(plot.title = element_text(hjust = 0.5), plot.subtitle = element_text(hjust = 0.5))

n1size_plot <- n1size_plot +
  theme(axis.title = element_text(size = 18)) +
  theme(plot.title = element_text(size = 20)) +
  theme(axis.text = element_text(size = 18)) +
  theme(legend.title = element_text(size = 18)) +
  theme(legend.text = element_text(size = 16)) +
  theme(plot.subtitle = element_text(size = 16))


d2size_plot <- sizebins %>%
  ggplot(aes(x = X , y = D2))+
  geom_bar(stat = 'identity', position = 'dodge', colour = "steelblue", fill = "steelblue")+
  labs(x = 'Body Length Size Classes',
       y = "Proportion of Individuals (%)",
       title = 'Site D2 in Trevor Channel',
       subtitle = '60m tow depth'
       )+
  theme(plot.title = element_text(hjust = 0.5), plot.subtitle = element_text(hjust = 0.5))

d2size_plot <- d2size_plot +
  theme(axis.title = element_text(size = 18)) +
  theme(plot.title = element_text(size = 20)) +
  theme(axis.text = element_text(size = 18)) +
  theme(legend.title = element_text(size = 18)) +
  theme(legend.text = element_text(size = 16)) +  
  theme(plot.subtitle = element_text(size = 16))


d1size_plot <- sizebins %>%
  ggplot(aes(x = X , y = D1))+
  geom_bar(stat = 'identity', position = 'dodge', colour = "steelblue", fill = "steelblue")+
  labs(x = 'Body Length Size Classes',
       y = "Proportion of Individuals (%)",
       title = 'Site D1 in Trevor Channel',
       subtitle = '120m tow depth'
       )+
  theme(plot.title = element_text(hjust = 0.5), plot.subtitle = element_text(hjust = 0.5))

d1size_plot <- d1size_plot +
  theme(axis.title = element_text(size = 18)) +
  theme(plot.title = element_text(size = 20)) +
  theme(axis.text = element_text(size = 18)) +
  theme(legend.title = element_text(size = 18)) +
  theme(legend.text = element_text(size = 16)) +
  theme(plot.subtitle = element_text(size = 16))

plot_comp <- srsize_plot + n1size_plot + d2size_plot + d1size_plot
plot_final <- plot_comp + plot_annotation(tag_levels = "A")

ggsave(here::here("size_plots2.png"), height = 9, width = 18) 

```


```{r}
###Biomass
biomasspercenttaxa <- read.csv("./data/biomass-percent2.csv", row.names = 1)
biomassvaluesgroup <- read.csv("./data/biomassvalues.csv", row.names = 1)
biomasspercentgroups <- read.csv("./data/biomasspercentgroups.csv", row.names = 1)

#Biomass percent by taxa


biomasspercentspecies <- tibble::rownames_to_column(biomasspercenttaxa, "Species")
#compositiongroups2$N1 <- as.double(compositiongroups2$N1)
biomasspercentspecies <- biomasspercentspecies %>%
  pivot_longer(!`Species`, names_to = "Site", values_to = "Count")


species_plot <- biomasspercentspecies %>%
  ggplot(aes(fill = fct_inorder(`Species`), x = Site, y = Count)) + 
    geom_bar(position="fill", stat="identity")+
  labs(x = "Site",
       y = "Proportion of Sample",
       fill = "Species name",
       title = "Zooplankton Biomass in Barkley Sound")+
  theme(plot.title = element_text(hjust = 0.5))+
  #scale_fill_excel_new("Office Theme")+
  #scale_fill_viridis(discrete = TRUE, option = "C")+
  scale_x_discrete(limits=rev)+
  scale_fill_manual(values = mycolours)

species_plot <- species_plot +
  theme(axis.title = element_text(size = 18)) +
  theme(plot.title = element_text(size = 20)) +
  theme(axis.text = element_text(size = 18)) +
  theme(legend.title = element_text(size = 18)) +
  theme(legend.text = element_text(size = 16))
print(species_plot)

ggsave(here::here("biomassspecies_plot2.png", height = 9, width = 14))

#Biomass stacked bar group

biomassvaluesgroup <- tibble::rownames_to_column(biomassvaluesgroup, "Taxon")
biomassvaluesgroup <- biomassvaluesgroup %>%
  pivot_longer(!`Taxon`, names_to = "Site", values_to = "Count")
#view(biomassvaluesgroup)


biomassvalues_plot <- biomassvaluesgroup %>%
  ggplot(aes(fill = `Taxon`, x = Site, y = Count)) + 
    geom_bar(stat = "identity")+
  labs(x = "Site",
       y = "Biomass (mgC/m^3)",
       fill = "Taxon name",
       title = "Zooplankton Biomass in Barkley Sound")+
  theme(plot.title = element_text(hjust = 0.5))+
  scale_x_discrete(limits=rev)+
  scale_fill_manual(values = mycolours)

biomassvalues_plot <- biomassvalues_plot +
  theme(axis.title = element_text(size = 18)) +
  theme(plot.title = element_text(size = 20)) +
  theme(axis.text = element_text(size = 18)) +
  theme(legend.title = element_text(size = 18)) +
  theme(legend.text = element_text(size = 16))

print(biomassvalues_plot)

ggsave(here::here("biomassvalues_plot2.png"), height = 9, width = 12)


#Biomass percent comp group
view(biomasspercentgroups)

biomasspercentgroups <- tibble::rownames_to_column(biomasspercentgroups, "Taxon")
biomasspercentgroups <- biomasspercentgroups %>%
  pivot_longer(!`Taxon`, names_to = "Site", values_to = "Count")


biomassgroups_plot <- biomasspercentgroups %>%
  ggplot(aes(fill = `Taxon`, x = Site, y = Count)) + 
    geom_bar(position="fill", stat="identity")+
  labs(x = "Site",
       y = "Proportion of Sample",
       fill = "Taxon name",
       title = "Zooplankton Biomass Composition in Barkley Sound")+
  theme(plot.title = element_text(hjust = 0.5))+
  scale_x_discrete(limits=rev)+
  scale_fill_manual(values = mycolours)

biomassgroups_plot <- biomassgroups_plot +
  theme(axis.title = element_text(size = 18)) +
  theme(plot.title = element_text(size = 20)) +
  theme(axis.text = element_text(size = 18)) +
  theme(legend.title = element_text(size = 18)) +
  theme(legend.text = element_text(size = 16))

biomassgroups_plot

ggsave(here::here("biomassgroups_plot.png"), height = 9, width = 12)


```


```{r}
#Diversity data to dataframe

#d1data below
d1counts <- read.csv("~/Desktop/d1counts.csv", row.names = 1)
d1div <- diversity(d1counts)
d1even <- d1div/(log(23))

#d2databelow
d2counts <- read.csv("~/Desktop/d2counts.csv", row.names = 1)
d2div <- diversity(d2counts)
d2even <- d2div/(log(19))

#n1databelow
n1counts <- read.csv("~/Desktop/n1counts.csv", row.names = 1)
n1div <- diversity(n1counts)
n1even <- n1div/(log(18))

#srdatabelow
srcounts <- read.csv("~/Desktop/srcounts.csv", row.names = 1)
srdiv <- diversity(srcounts)
sreven <- srdiv/(log(17))

#make dataframe and rename columns

a <- list(srdiv, n1div, d2div, d1div)
b <- list(sreven, n1even, d2even, d1even)

df <- tibble(a, b)
colnames1 <- c("Shannon-Weaver Diversity Index", "Pielou's Species Evenness")
rownames1 <- c("SR", "N1", "D2", "D1")
colnames(df) <- colnames1
rownames(df) <- rownames1

```

```{r}
##Biomass cont by size class

sizebiomass <- read.csv("./data/sizebiomass.csv")

srsize_plot <- sizebiomass %>%
  ggplot(aes(x = fct_inorder(X) , y = SR))+
  geom_bar(stat = 'identity', position = 'dodge', colour = "steelblue", fill = "steelblue")+
  labs(x = 'Biomass Contribution by Size Class',
       y = "Proportion of Individuals (%)",
       title = 'Site SR1 at Sarita River',
       subtitle = '30m tow depth'
       )+
    ylim(0,50)+
  theme(plot.title = element_text(hjust = 0.5), plot.subtitle = element_text(hjust = 0.5))

srsize_plot <- srsize_plot +
  theme(axis.title = element_text(size = 18)) +
  theme(plot.title = element_text(size = 20)) +
  theme(axis.text = element_text(size = 18)) +
  theme(legend.title = element_text(size = 18)) +
  theme(legend.text = element_text(size = 16)) +
  theme(plot.subtitle = element_text(size = 16))
print(srsize_plot)

n1size_plot <- sizebiomass %>%
  ggplot(aes(x = fct_inorder(X) , y = N1))+
  geom_bar(stat = 'identity', position = 'dodge', colour = "steelblue", fill = "steelblue")+
  labs(x = 'Biomass Contribution by Size Class',
       y = "Proportion of Individuals (%)",
       title = 'Site N1 in Trevor Channel',
       subtitle = '30m tow depth'
       )+
    ylim(0,50)+
  theme(plot.title = element_text(hjust = 0.5), plot.subtitle = element_text(hjust = 0.5))

n1size_plot <- n1size_plot +
  theme(axis.title = element_text(size = 18)) +
  theme(plot.title = element_text(size = 20)) +
  theme(axis.text = element_text(size = 18)) +
  theme(legend.title = element_text(size = 18)) +
  theme(legend.text = element_text(size = 16)) +
  theme(plot.subtitle = element_text(size = 16))


d2size_plot <- sizebiomass %>%
  ggplot(aes(x = fct_inorder(X) , y = D2))+
  geom_bar(stat = 'identity', position = 'dodge', colour = "steelblue", fill = "steelblue")+
  labs(x = 'Biomass Contribution by Size Class',
       y = "Proportion of Individuals (%)",
       title = 'Site D2 in Trevor Channel',
       subtitle = '60m tow depth'
       )+
    ylim(0,50)+
  theme(plot.title = element_text(hjust = 0.5), plot.subtitle = element_text(hjust = 0.5))

d2size_plot <- d2size_plot +
  theme(axis.title = element_text(size = 18)) +
  theme(plot.title = element_text(size = 20)) +
  theme(axis.text = element_text(size = 18)) +
  theme(legend.title = element_text(size = 18)) +
  theme(legend.text = element_text(size = 16)) +  
  theme(plot.subtitle = element_text(size = 16))


d1size_plot <- sizebiomass %>%
  ggplot(aes(x = fct_inorder(X) , y = D1))+
  geom_bar(stat = 'identity', position = 'dodge', colour = "steelblue", fill = "steelblue")+
  labs(x = 'Biomass Contribution by Size Class',
       y = "Proportion of Individuals (%)",
       title = 'Site D1 in Trevor Channel',
       subtitle = '120m tow depth'
       )+
  theme(plot.title = element_text(hjust = 0.5), plot.subtitle = element_text(hjust = 0.5))

d1size_plot <- d1size_plot +
  theme(axis.title = element_text(size = 18)) +
  theme(plot.title = element_text(size = 20)) +
  theme(axis.text = element_text(size = 18)) +
  theme(legend.title = element_text(size = 18)) +
  theme(legend.text = element_text(size = 16)) +
  theme(plot.subtitle = element_text(size = 16))

plot_comp <- srsize_plot + n1size_plot + d2size_plot + d1size_plot
plot_final <- plot_comp + plot_annotation(tag_levels = "A")
print(plot_final)

ggsave("size_biomass.png", height = 9, width = 18) 


```

```{r}
#Map & bathymetry stuff
# get bathymetry data
b = getNOAA.bathy(lon1 = -125.2, lon2 = -124.1167, lat1 = 48.95, lat2 = 48.70, 
                  resolution = 1)
## Querying NOAA database ...
## This may take seconds to minutes, depending on grid size
## Building bathy matrix ...

# make a simple track line
#lin = data.frame(
  #lon = c(-65.17536, -65.37423, -65.64541, -66.06122, -66.15161),  
  #lat = c(43.30837, 42.94679, 42.87448, 42.92871, 42.72985))

# make a few points
pts = data.frame(
  lon = c(-125.2639, -125.2542, -125.4222, -125.0842),
  lat = c(49.09167, 49.06833, 48.89778, 49.1375)
)

data("coastlineWorldFine")

# convert bathymetry
bathyLon = as.numeric(rownames(b))
bathyLat = as.numeric(colnames(b))
bathyZ = as.numeric(b)
dim(bathyZ) = dim(b)

# define plotting region
mlon = mean(pts$lon)
mlat = mean(pts$lat)
span = 70

# plot coastline (no projection)
plot(coastlineWorldFine, clon = mlon, clat = mlat, span = span)

# plot bathymetry
contour(bathyLon,bathyLat,bathyZ,
        levels = c(-50, -100, -150, -200, -250),
        lwd = c(1, 1, 2, 2, 3),
        lty = c(3, 1, 3, 1, 3),
        drawlabels = F, add = TRUE, col = 'darkgray')

# add depth legend
legend("bottomright", seg.len = 3, cex = 0.8,
       lwd = c(1, 1, 2, 2, 3),
       lty = c(3, 1, 3, 1, 3),
       legend = c("50", "100", "150", "200", "250"),
       col = 'darkgray', title = "Depth [m]", bg= "white")

# add map data
points(pts, pch = 16, col = 'red')
lines(lin, col = 'blue')
polygon(ply, lty = 2)

```

```{r}
#bathy cont
dt <- data.frame(lat = c(48, 48, 50, 50), lon = c(-124, -125, -125, -124))
pts = data.frame(lon = c(125.2639, 125.2542, 125.4222, 125.0842), lat = c(49.09167, 49.06833, 48.89778, 49.1375))
basemap(data = dt, bathymetry = TRUE)+
  geom_point(data = pts, aes(x = lon, y = lat))



```